### 2025-06-29 paper/quarto_aaai/paper.tex
Fix overview of CT

\begin{enumerate}
    \item (Faithfulness) $\int^{A} p_\theta(\mathbf{x}^\prime|\mathbf{y}^{+})d\mathbf{x} \rightarrow 1$, where $A$ is some arbitrarily small region around $\mathbf{x}^{\prime}$.
    \item (Plausibility) $\int^{A} p(\mathbf{x}^\prime|\mathbf{y}^{+})d\mathbf{x} \rightarrow 1$; $A$ as specified above.
    \item (Actionability) Permutations $\Delta$ are subject to some actionability constraints.
\end{enumerate}
and $p_\theta(\mathbf{x}|\mathbf{y}^{+})$ denotes the conditional posterior distribution over inputs. For simplicity, we refer to a model with high explanatory capacity as \textbf{explainable} in this manuscript. 
\end{definition}

@@

Let \(\mathbf{x}_t^\prime\) for \(t=0,...,T\) denote a counterfactual
generated through gradient descent over \(T\) iterations as originally
proposed by \citet{wachter2017counterfactual}. In broad terms,
generating CEs using gradient descent entails solving some form of an
objective that balances (1) the classification loss
\(\text{yloss}(\mathbf{M}_\theta(\mathbf{x}),\mathbf{y})\), and (2) one
or more penalty terms \(\lambda_{i}\text{cost}_{i}(\cdot)\). The exact
specification of these penalties induces various properties in the
counterfactual outcomes, and tends to be the key feature that
distinguishes various gradient-based ``generators'' or ``explainers''
(CE algorithms) from the literature \citep{altmeyer2023explaining}.

@@

CT leverages gradient-based CE search already in training to produce
perturbed versions of the factual samples \(\mathbf{x}\). Initially,
until \(\tau\) is reached or when \(t\leq T\), generated samples
\(\mathbf{x}_t^\prime\) serve as adversarial examples; we term them
\emph{nascent} counterfactuals \(\mathbf{x}_{t\leq T}^\prime\). Over
multiple training iterations up to time \(T\), these nascent
counterfactuals move towards the decision boundary and turn into
\emph{mature} counterfactuals \(\mathbf{x}_T^\prime\) when they cross
into the target class or when \(t > T\).

>>>>

\begin{enumerate}
    \item (Faithfulness) $\int^{A} p_\theta(\mathbf{x}^\prime|\mathbf{y}^{+})d\mathbf{x} \rightarrow 1$, where $A$ is some arbitrarily small region around $\mathbf{x}^{\prime}$.
    \item (Plausibility) $\int^{A} p(\mathbf{x}^\prime|\mathbf{y}^{+})d\mathbf{x} \rightarrow 1$; $A$ as specified above.
    \item (Actionability) Perturbations $\Delta$ are subject to some actionability constraints.
\end{enumerate}
and $p_\theta(\mathbf{x}|\mathbf{y}^{+})$ denotes the conditional posterior distribution over inputs. For simplicity, we refer to a model with high explanatory capacity as \textbf{explainable} in this manuscript. 
\end{definition}

@@

Let \(\mathbf{x}_t^\prime\) for \(t=0,...,T\) denote a counterfactual
generated through gradient descent over \(T\) iterations as originally
proposed by \citet{wachter2017counterfactual}. In broad terms, searching
for CEs using gradient descent entails optimizing some form of an
objective that balances (1) the classification loss
\(\text{yloss}(\mathbf{M}_\theta(\mathbf{x}),\mathbf{y})\), and (2) one
or more penalty terms \(\lambda_{i}\text{cost}_{i}(\cdot)\). The exact
specification of these penalties induces various properties in the
counterfactual outcomes, and tends to be the key feature that
distinguishes various gradient-based ``generators'' or ``explainers''
(CE algorithms) from the literature \citep{altmeyer2023explaining}.

@@

CT adapts gradient-based CE search already in training to generate
\(\mathbf{x}^\prime\), the perturbed versions of the factual samples
\(\mathbf{x}\). We use the term \emph{nascent} to denote counterfactuals
for which the search has not converged, \(\mathbf{x}_{t\leq T}^\prime\).
Initially, when the perturbations are small, nascent counterfactuals
serve as AEs. Over multiple iterations, they move towards the decision
boundary and turn into \emph{mature} counterfactuals
\(\mathbf{x}_T^\prime\) when they cross into the target class (becoming
CEs) or when \(t > T\).

### 2025-06-29 paper/sections/method.qmd
Fix overview of CT

Let $\mathbf{x}_t^\prime$ for $t=0,...,T$ denote a counterfactual generated through gradient descent over $T$ iterations as originally proposed by @wachter2017counterfactual. In broad terms, generating CEs using gradient descent entails solving some form of an objective that balances (1) the classification loss $\text{yloss}(\mathbf{M}_\theta(\mathbf{x}),\mathbf{y})$, and (2) one or more penalty terms $\lambda_{i}\text{cost}_{i}(\cdot)$. The exact specification of these penalties induces various properties in the counterfactual outcomes, and tends to be the key feature that distinguishes various gradient-based "generators" or "explainers" (CE algorithms) from the literature [@altmeyer2023explaining].

@@

CT leverages gradient-based CE search already in training to produce perturbed versions of the factual samples $\mathbf{x}$. Initially, until $\tau$ is reached or when $t\leq T$, generated samples $\mathbf{x}_t^\prime$ serve as adversarial examples; we term them *nascent* counterfactuals $\mathbf{x}_{t\leq T}^\prime$. Over multiple training iterations up to time $T$, these nascent counterfactuals move towards the decision boundary and turn into *mature* counterfactuals $\mathbf{x}_T^\prime$ when they cross into the target class or when $t > T$.

>>>>

Let $\mathbf{x}_t^\prime$ for $t=0,...,T$ denote a counterfactual generated through gradient descent over $T$ iterations as originally proposed by @wachter2017counterfactual. In broad terms, searching for CEs using gradient descent entails optimizing some form of an objective that balances (1) the classification loss $\text{yloss}(\mathbf{M}_\theta(\mathbf{x}),\mathbf{y})$, and (2) one or more penalty terms $\lambda_{i}\text{cost}_{i}(\cdot)$. The exact specification of these penalties induces various properties in the counterfactual outcomes, and tends to be the key feature that distinguishes various gradient-based "generators" or "explainers" (CE algorithms) from the literature [@altmeyer2023explaining].

@@

CT adapts gradient-based CE search already in training to generate $\mathbf{x}^\prime$, the perturbed versions of the factual samples $\mathbf{x}$. We use the term *nascent* to denote counterfactuals for which the search has not converged, $\mathbf{x}_{t\leq T}^\prime$. Initially, when the perturbations are small, nascent counterfactuals serve as AEs. Over multiple iterations, they move towards the decision boundary and turn into *mature* counterfactuals $\mathbf{x}_T^\prime$ when they cross into the target class (becoming CEs) or when $t > T$.
